---
title: "Bibliometric Analysis-Eval Data Viz"
format: html
editor: visual
data: 03-24-2025 
footer: "Data Visualisation and Communication in Evaluation (Richard Amoako)"
---

**Title: Data Visualization and Communication in Evaluation

```{r message=FALSE, warning=FALSE, echo=FALSE}
#| echo: false
#| eval: false
#| include: false
#| message: false
#| warning: false
#| error: false
#| results: false
#| setup: false
#| import: false
#| options: false
#| code_folding: false
#| cache: false
```

```{r}
#Load packages 
library(tidyverse)
library(janitor)
library(psych)
library(ggplot2)
library(dplyr)

```


This study employs bibliometric methods using the PRISMA framework to address critical questions about data visualization evolution, key influences, thematic foci, and global distribution of research on data visualization within the evaluation field. Through a comprehensive analysis of scholarly publications, this research aims to provide a valuable resource for researchers, practitioners, and policymakers seeking to understand and leverage the power of data visualization to improve evaluation practice.

*Research Questions*

RQ1- How has data visualization in evaluation evolved over time?




*Research Design*

This study employs a bibliometric analysis approach to comprehensively map the intellectual landscape of data visualization in evaluation research. The methodology is complemented by a systematic review process, adhering to the Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) guidelines, to ensure a structured, transparent, and replicable selection of relevant literature.


*Bibliometric Analysis:* The study applies a range of bibliometric techniques, including: Co-citation analysis: to identify clusters of highly cited publications and reveal the intellectual foundations of the field. Keyword co-occurrence mapping: exploring thematic trends and emerging research areas by analyzing the relationships between keywords used in the literature. Collaboration network analysis: examines patterns of collaboration among authors, institutions, and countries, and identifies key actors and knowledge hubs.





###Below is analysis and output of the data extraction and analysis

```{{r}}
```

<div>

| Database                        | Number  |
|---------------------------------|---------|
| Scopus                          | 950     |
| WoS                             | 552     |
| Dupblicates                     | \(88\)  |
| Removed papers dated below 2010 | \(147\) |
| Total                           | 1267    |

: Data Extraction

</div>

```{r}
#Load data
biblio_df <- read_csv("biblio_data.csv")

##Check the column name of DOI to doi
biblio_df <- biblio_df %>% rename(doi = DOI)

```

*Data Preprocessing*

**Removing Duplicates**

Total data extracted from the two databases is 1505.

```{r}
#calculate number of duplicated rows
sum(duplicated(biblio_df$doi))
```

```{r}
#remove duplicates by matching doi
new_biblio <- biblio_df %>% distinct(doi, .keep_all = TRUE)

#check for duplicates
sum(duplicated(new_biblio$doi))
```

**Screening data by title and abstract**

```{r}
#Screen title and abstract for irrelavant data
new_biblio <- new_biblio %>% filter(!str_detect(title, "^(?i)abstract"))

#Additional screening is being done using AsReview Lab
```

##Full text screening Full text screening was performed using Loon Lens.

**Data Cleaning**

```{r}

#check the first few rows of the data
head(new_biblio)

```

```{r}
#Revised inclusion criteria: Time range: (2010–2025)
#filter by years 2010-2025 and create a new dataframe called rnew_biblio
rnew_biblio <- new_biblio %>% 
  filter(year >= 2010 & year <= 2025)

#Clean column names
rnew_biblio <- rnew_biblio %>% 
  clean_names()
```

```{r}
# Handle missing data (NA)
# Check for missing data
missing_summary <- rnew_biblio %>%
  summarise(across(everything(), ~ sum(is.na(.))))

#print (missing_summary)
```

**Processed Data**

```{r}
#save new biblio
write_csv(rnew_biblio, "rnew_biblio.csv")
```

. . .

```         
                     NEW UPDATES 03/24/2025
       
       
        **Data Screening and Preliminary Data Visualization**
        
```

##Full text screening Full text screening was performed using Loon Lens. The screened data was further cleaned and used for the Data visualization. Total of 326 out of 1267 met the inclusion criteria

```{r}
#Load new data merged_data.csv
df_merged <- read_csv("merged_data.csv")

#clean column names
df_merged <- df_merged %>% clean_names()

#check the structure of the data
#str(df_merged)

#check the first few rows of the data
head(df_merged)
```

```{r}
#Preprocess the data
#set the text variables as factors
df_merged <- df_merged %>%
  mutate(across(where(is.character), as.factor))

#Convert these variables volume, issue, page_count, cite_count to numeric
df_merged <- df_merged %>%
  mutate(across(c(volume, issue, page_count, cite_count), as.numeric))
```

**Additional Data Cleaning**

```{r}
#filter to remove paper with low confidence using -Low
df_merged <- df_merged %>%
  filter(!str_detect(confidence, "Low"))

#filter to remove paper with paper type -conference paper
df_merged <- df_merged %>%
  filter(!str_detect(paper_type, "Conference paper"))
```

```{r}
#Check for missing data
missin <- df_merged %>%
  summarise(across(everything(), ~ sum(is.na(.))))
#print (missin)

```

**Preliminary Analysis**

```{r}
#Summary statistics
summary(df_merged$cite_count)

```

```{r}
#Descriptive statistics
# Frequency table with counts and percentages for gender
freq_year <- df_merged %>%
  tabyl(year) %>%  # Create frequency table
  adorn_pct_formatting()  # Format percentages

freq_confi <- df_merged %>%
  tabyl(confidence) %>%  # Create frequency table
  adorn_pct_formatting()  # Format percentages

freq_country <- df_merged %>%
  tabyl(country) %>%  # Create frequency table
  adorn_pct_formatting()  # Format percentages

freq_paper <- df_merged %>%
  tabyl(paper_type) %>%  # Create frequency table
  adorn_pct_formatting()  # Format percentages

# Print the frequency table
print(freq_year)
print(freq_confi)
#print(freq_country)
print(freq_paper)

```

```{r}
#filter papers no citation count
df_mergedx <- df_merged %>%
  filter(cite_count > 0)
```

```{r}
#Visualize the distribution of citation counts
library(ggplot2)

stopifnot(
  "df_mergedx not found" = exists("df_mergedx"),
  "cite_count column missing" = "cite_count" %in% colnames(df_mergedx),
  "cite_count contains non-numeric values" = is.numeric(df_mergedx$cite_count)
)


df_mergedx %>%
   filter(!is.na(cite_count)) %>%
  ggplot(aes(x = cite_count)) +
  geom_histogram(binwidth = 10, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Citation Counts",
       x = "Citation Count",
       y = "Frequency") +
  theme_minimal()
```

```{r}
#Conduct a citation analysis to identify influential papers
#Identify key authors and institutions
#Conduct a keyword co-occurrence analysis
#Conduct a network analysis to identify collaboration patterns
#Conduct a content analysis of the visualization techniques used

```

```{r}
# #Load bibliometrix package
# library(bibliometrix)
# 
# #Create a bibliographic data frame
# #biblio <- convert2df("df_merge.csv")
# data <- convert2df("scopus_data.csv", dbsource = "scopus", format = "csv")
# summary(data)
```

**Trend Analysis**

```{r}
# 1. Trend Analysis: Annual publication counts
yearly_pubs <- df_mergedx %>%
  group_by(year) %>%
  summarise(count = n()) %>%
  arrange(year)

# Plot annual publication counts
ggplot(yearly_pubs, aes(x = year, y = count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_line(color = "red", size = 1) +
  geom_point(color = "red", size = 3) +
  theme_minimal() +
  labs(title = "Annual Publication Counts",
       x = "Year",
       y = "Number of Publications") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
library(ggplot2)
p <- ggplot(mtcars, aes(mpg, hp)) + geom_point()

# Save to working directory
#ggsave("my_ggplot.png", plot = p, width = 8, height = 6, dpi = 300)
```

```{r}
# Calculate yearly average citations
yearly_citations <- df_mergedx %>%
  group_by(year) %>%
  summarise(
    total_citations = sum(cite_count, na.rm = TRUE),
    avg_citations = mean(cite_count, na.rm = TRUE),
    count = n()
  ) %>%
  arrange(year)

# Plot yearly average citations
ggplot(yearly_citations, aes(x = year)) +
  geom_bar(aes(y = total_citations), stat = "identity", fill = "skyblue", alpha = 0.7) +
  geom_line(aes(y = avg_citations * max(total_citations)/max(avg_citations)), 
            color = "red", linewidth = 1) +
  geom_point(aes(y = avg_citations * max(total_citations)/max(avg_citations)), 
             color = "red", size = 3) +
  scale_y_continuous(
    name = "Total Citations",
    sec.axis = sec_axis(~. * max(yearly_citations$avg_citations)/max(yearly_citations$total_citations), 
                        name = "Average Citations per Paper")
  ) +
  theme_minimal() +
  labs(title = "Yearly Citation Patterns",
       x = "Year") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.y.left = element_text(color = "skyblue4"),
        axis.title.y.right = element_text(color = "red"))
```

```{r}
# 2. Citation and Impact Analysis
# Identify top-cited papers
top_papers <- df_mergedx %>%
  arrange(desc(cite_count)) %>%
  select(title, first_author, journal, year, cite_count) %>%
  head(10)

# Display top-cited papers
print(top_papers)

# Identify high-impact journals
journal_impact <- df_mergedx %>%
  group_by(journal) %>%
  summarise(
    papers = n(),
    total_citations = sum(cite_count, na.rm = TRUE),
    avg_citations = mean(cite_count, na.rm = TRUE)
  ) %>%
  filter(papers >= 3) %>%  # Only include journals with at least 3 papers
  arrange(desc(total_citations))

# Display top journals by total citations
print(head(journal_impact, 10))
```

```{r}

# Load world map data
#world_map <- map_data("world")

# # World map to show the distribution of publications by country
# df_merged %>%
#   count(country) %>%
#   ggplot(aes(fill = n, map_id = country)) +
#   geom_map(map = world_map, aes(map_id = countries), color = "white", data = world_map) +
#   expand_limits(x = world_map$long, y = world_map$lat) +  # Ensure the map fits within the plot limits
#   scale_fill_viridis_c() +
#   labs(title = "Distribution of Publications by Country",
#        fill = "Number of Publications") +
#   theme_void()
```

*##Next Steps*

#Continue bilbiometric analysis

#Create more visualizations to represent the data

-   Identify key authors and institutions

-   Conduct a keyword co-occurrence analysis

-   Conduct a network analysis to identify collaboration patterns

-   Conduct a topic modeling on key themes or main research focus of the studies

------------------------------------------------------------------------

```         
         **Week 12: Progress Updates 5**04/07/2025
```

I filtered the data to exclude papers published in 2025 because these papers have no citations. The include criteria states papers should have at least one citation. I also filtered out papers with low confidence. The data was further cleaned to remove papers with no citation counts.

```{r}
# #filter papers no citation count
# df_mergedx <- df_merged %>%
#   filter(cite_count > 0)

#check the number of rows after removing irrelevant data
nrow(df_mergedx)

#check the first few rows of the data
head(df_mergedx)

#filter papers with low confidence
df_mergedx <- df_mergedx %>%
  filter(!str_detect(confidence, "Low"))

#save new df_mergedx
write_csv(df_mergedx, "df_mergedx.csv")

```

```{r}
# Load necessary libraries
#library(dplyr)
#library(ggplot2)
#library(stringr)
#library(tidyr)

###Topic Authors and Affiliations


# Get the top authors
top_authors <- df_mergedx %>%
  count(first_author) %>%
  filter(!is.na(first_author) & first_author != "") %>%
  arrange(desc(n)) %>%
  head(10)

print(top_authors)
```

```{r}
# Process affiliations: Some rows have multiple institutions separated by semicolons or commas.
# I first try to split by semicolon.

aff_df <- df_mergedx %>% 
  filter(!is.na(affiliation) & affiliation != "") %>% 
  mutate(affiliation = as.character(affiliation)) %>%  # To ensure affiliation is character
  mutate(affiliation_list = strsplit(affiliation, ";")) %>% 
  unnest(affiliation_list) %>% 
  mutate(affiliation_clean = trimws(affiliation_list))

# Count frequency of each affiliation
top_affiliations <- aff_df %>%
  count(affiliation_clean, sort = TRUE) %>%
  head(10)

print(top_affiliations)

# Plot the top affiliations with ggplot2 with a clear white background
plot_affiliations <- ggplot(top_affiliations, aes(x = reorder(affiliation_clean, n), y = n)) +
  geom_bar(stat = 'identity', fill = 'steelblue') +
  coord_flip() +
  labs(title = 'Top Affiliations', x = 'Affiliation', y = 'Count') +
  theme_minimal(base_size = 12) +
  theme(panel.background = element_rect(fill = 'white'))

print(plot_affiliations)

# Additionally, visualize top authors (already computed before) similarly
plot_authors <- ggplot(top_authors, aes(x = reorder(first_author, n), y = n)) +
  geom_bar(stat = 'identity', fill = 'tomato') +
  coord_flip() +
  labs(title = 'Top Authors', x = 'Author', y = 'Count') +
  theme_minimal(base_size = 12) +
  theme(panel.background = element_rect(fill = 'white'))

print(plot_authors)
```

**Mapping publication count by Country**

```{r}
#library(dplyr)
#library(ggplot2)
#library(maps)

# Count the number of publications per country
country_counts <- df_mergedx %>% 
  filter(!is.na(country) & country != "") %>% 
  count(country, sort = TRUE)

# Prepare world map data
world_map <- map_data("world")

# Harmonize country names if necessary (not shown here)

# Join the counts with the world map data
country_counts <- country_counts %>% rename(region = country)
map_data_counts <- left_join(world_map, country_counts, by = "region")

# Create the map plot
map_plot <- ggplot(map_data_counts, aes(x = long, y = lat, group = group, fill = n)) +
  geom_polygon(color = "gray90", size = 0.1) +
  scale_fill_gradient(
    name = "Publication Count",
    low = "lightblue",
    high = "darkblue",
    na.value = "white"
  ) +
  theme_minimal() +
  labs(title = "Geographical Distribution of Publications") +
  theme(
    panel.background = element_rect(fill = "white"),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank()
  )

print(map_plot)
```

```{r}

# #library(viridis)
# #library(rnaturalearth)
# world <- ne_countries(scale = "medium", returnclass = "sf")
# merged_data <- merge(world, df_mergedx, by.x = "iso_a3", by.y = "ISO Code")

map_plot1 <- ggplot(map_data_counts, aes(x = long, y = lat, group = group, fill = n)) +
  geom_polygon(color = "gray90", size = 0.1) +
  scale_fill_viridis_c(
    name = "Publication Count",
    na.value = "white",
    option = "D"
  ) +
  theme_minimal() +
  labs(title = "Geographical Distribution of Publications") +
  theme(panel.background = element_rect(fill = "white"))
print (map_plot1)
```

```{r}
library(sf)
library(rnaturalearth)
library(dplyr)

# Your publication data
pub_data <- data.frame(
  country = c("United States", "United Kingdom", "Germany", "Australia", 
              "Canada", "Italy", "Netherlands", "Spain", "Brazil", "Hong Kong"),
  count = c(98, 31, 13, 12, 12, 12, 8, 8, 5, 5)
)

# Standardize country names
pub_data <- pub_data %>%
  mutate(country = case_when(
    country == "United States" ~ "United States of America",
    country == "Hong Kong" ~ "Hong Kong S.A.R.",
    TRUE ~ country
  ))

# Load world map and merge
world <- ne_countries(scale = "medium", returnclass = "sf")
merged_data <- world %>% 
  left_join(pub_data, by = c("name" = "country"))

# Convert to 'sf' (redundant here since 'world' is already 'sf', but safe)
merged_sf <- st_as_sf(merged_data)  # Now works!

# Plot with ggplot2
ggplot(merged_sf) +
  geom_sf(aes(fill = count)) +
  scale_fill_viridis_c(na.value = "grey90")
```

```{r}
# library(sf)
# library(rnaturalearth)
# library(dplyr)
# library(ggplot2)
# library(viridis)

# Your publication counts data
pub_data <- data.frame(
  country = c("United States", "United Kingdom", "Germany", "Australia", 
              "Canada", "Italy", "Netherlands", "Spain", "Brazil", "Hong Kong",
              "Indonesia", "Chile", "China", "Denmark", "Japan", "Malaysia", 
              "South Africa", "Sweden", "France", "India", "Bangladesh", 
              "Ireland", "New Zealand", "Norway", "Switzerland", "Taiwan", 
              "Uruguay", "Belgium", "Burkina Faso", "Croatia", "Czech Republic",
              "Ecuador", "Egypt", "Ethiopia", "Finland", "Kenya", "Lebanon", 
              "Malawi", "Morocco", "Nigeria", "Poland", "Portugal", "Singapore",
              "Slovenia", "South Korea", "Sri Lanka", "Uganda", "Viet Nam"),
  count = c(98, 31, 13, 12, 12, 12, 8, 8, 5, 5, 5, 4, 4, 4, 4, 4, 4, 4, 3, 3,
            2, 2, 2, 2, 2, 2, 2, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
            1, 1, 1, 1, 1, 1)
)

# Standardize country names to match 'rnaturalearth'
pub_data <- pub_data %>%
  mutate(country = case_when(
    country == "United States" ~ "United States of America",
    country == "Hong Kong" ~ "Hong Kong S.A.R.",
    country == "Czech Republic" ~ "Czechia",
    country == "South Korea" ~ "Republic of Korea",
    country == "Viet Nam" ~ "Vietnam",
    country == "Taiwan" ~ "Taiwan Province of China",  # Note: Map may not display due to geopolitical constraints
    TRUE ~ country
  ))

# Load world map (medium resolution) and merge
world <- ne_countries(scale = "medium", returnclass = "sf") %>%
  select(name, iso_a3, geometry)

merged_data <- world %>%
  left_join(pub_data, by = c("name" = "country"))

# Plot
pubmap <- ggplot(merged_data) +
  geom_sf(aes(fill = count), color = "white", size = 0.2) +
  scale_fill_viridis(
    name = "Publications",
    option = "plasma",
    na.value = "grey90",
    limits = c(0, 100),
    breaks = c(0, 10, 30, 50, 75, 100)
  ) +
  labs(
    title = "Academic Publications by Country",
    caption = paste("Highest count: USA (98) |", 
                    "Total countries:", nrow(pub_data))
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    legend.position = "bottom"
  )

print (pubmap)
# Save the plot (optional)
ggsave("pub_map.jpg", width = 10, height = 6, dpi = 300)
```

```{r}
# library(leaflet)
# library(leaflet.extras)

merged_sf <- st_as_sf(merged_data)

leaflet(merged_sf) %>%
  addPolygons(
    fillColor = ~colorBin("YlOrRd", count)(count),
    fillOpacity = 0.7,
    color = "white",
    weight = 1,
    label = ~paste(name, ": ", count, " publications")
  ) %>%
  addLegend(
    pal = colorBin("YlOrRd", merged_sf$count, bins = 5),
    values = ~count,
    title = "Publications"
  )
```

```{r}
library(htmlwidgets)  # For saving as HTML

# Load world map and merge
world <- ne_countries(scale = "medium", returnclass = "sf")
merged_data <- world %>% 
  left_join(pub_data, by = c("name" = "country"))

# Create interactive map
map <- leaflet(merged_data) %>%
  addPolygons(
    fillColor = ~colorBin("YlOrRd", count, bins = 5)(count),
    fillOpacity = 0.7,
    color = "white",
    weight = 1,
    label = ~paste(name, ":", count, "publications"),
    highlightOptions = highlightOptions(
      weight = 2,
      color = "#666",
      fillOpacity = 0.9,
      bringToFront = TRUE
    )
  ) %>%
  addLegend(
    pal = colorBin("YlOrRd", merged_data$count, bins = 5),
    values = ~count,
    title = "Publications",
    position = "bottomright"
  ) %>%
  addProviderTiles(providers$CartoDB.Positron)  # Clean base map

# Save as HTML
saveWidget(map, file = "publications_map.html", selfcontained = TRUE)
```

**Keyword Co-occurrence Analysis**

```{r}
# keyword analysis using base igraph for visualization library(dplyr)
library(stringr)
library(igraph)

# I am working with the author_keywords column

# Filter rows with non-empty keywords
keywords_df <- df_mergedx %>% filter(!is.na(author_keywords) & author_keywords != "") %>%
  select(id, author_keywords)

# Tokenize and clean keywords: They are separated by semicolons
# I'll first replace semicolons with commas then split by comma.
keywords_df <- keywords_df %>%
  mutate(author_keywords = str_replace_all(author_keywords, ";", ",")) %>%
  mutate(keyword_list = strsplit(author_keywords, ","))

# Unnest the keyword_list
tokenized_keywords <- keywords_df %>%
  unnest(keyword_list) %>%
  mutate(keyword = tolower(trimws(keyword_list))) %>%
  filter(keyword != "")

# Count keyword frequencies
keyword_counts <- tokenized_keywords %>%
  count(keyword, sort = TRUE) %>%
  filter(n >= 3)  # Filter for keywords that appear at least 3 times

print(head(keyword_counts, 10))
```

```{r}
# To compute co-occurrence counts using tokenized keywords
# First, to get unique keywords for each record so that repeated keywords in one record don't result double counting
unique_keywords <- tokenized_keywords %>% group_by(id) %>% summarise(keywords = list(unique(keyword)))

# Function to generate all pairs of unique keywords for a record (if count>1)
cooccur_pairs <- lapply(unique_keywords$keywords, function(x) {
  if(length(x) > 1) {
    pairs <- t(combn(sort(x), 2))
    as.data.frame(pairs, stringsAsFactors = FALSE)
  } else {
    NULL
  }
})

# Bind all pairs
cooccurrence_df <- do.call(rbind, cooccur_pairs)
if(!is.null(cooccurrence_df) && nrow(cooccurrence_df) > 0) {
  names(cooccurrence_df) <- c('keyword1', 'keyword2')
  # Count the frequency of each co-occurring pair
  cooccurrence_counts <- cooccurrence_df %>% 
    group_by(keyword1, keyword2) %>% 
    summarise(n = n(), .groups = 'drop') %>% 
    arrange(desc(n))
} else {
  cooccurrence_counts <- data.frame(keyword1=character(), keyword2=character(), n=integer())
}

print(head(cooccurrence_counts, 10))

# Build an igraph network with a filter: only include co-occurrence pairs that occur at least 2 times
min_occurrence <- 2
edge_list <- cooccurrence_counts %>% filter(n >= min_occurrence)

if(nrow(edge_list) > 0) {
  g <- graph_from_data_frame(edge_list, directed = FALSE)
  
  # Basic plot using igraph's plot function
  set.seed(42)
  plot(g, edge.width = E(g)$n, vertex.label.cex = 0.8, main = 'Keyword Co-occurrence Network')
} else {
  print('Not enough co-occurrence data to build network.')
}
```

**Network Analysis of collaborations**

```{r}
# Check head of the 'first_author' and 'co_authors' columns
#head(df[, c('first_author', 'co_authors')])

# Filter out rows with non-empty co_authors column
collab_df <- df_mergedx %>% filter(!is.na(co_authors) & co_authors != "") %>%
  select(first_author, co_authors)

# Tokenize the co_authors into individual authors; assuming they are separated by commas or semicolons.
collab_df <- collab_df %>%
  mutate(co_authors = str_replace_all(co_authors, ";", ",")) %>%
  mutate(co_author_list = strsplit(co_authors, ","))

# Clean co_author names: trim and lowercase maybe
collab_df <- collab_df %>%
  rowwise() %>%
  mutate(co_author_list = list(trimws(co_author_list))) %>%
  ungroup()

# Build the edge list: each edge from first_author to each co-author
edge_list <- collab_df %>%
  select(first_author, co_author_list) %>%
  tidyr::unnest(co_author_list) %>%
  filter(co_author_list != "") %>%
  rename(co_author = co_author_list)

# Remove any self-loops if first_author equals co_author
edge_list <- edge_list %>% filter(first_author != co_author)

# Create a graph using igraph (undirected)
g <- graph_from_data_frame(edge_list, directed = FALSE)

# Compute network metrics: degree centrality, closeness, betweenness
deg <- degree(g, mode = 'all')
clos <- closeness(g, mode = 'all')
between <- betweenness(g, directed = FALSE)

# Add the metrics as vertex attributes
V(g)$degree <- deg
V(g)$closeness <- clos
V(g)$betweenness <- between

# Print summary of metrics for top 10 vertices by degree
top_deg <- sort(deg, decreasing = TRUE)[1:10]
print(top_deg)

# Visualize the network graph; adjust vertex size proportional to degree
set.seed(123)
plot(g, vertex.size = 5 + 2*deg, vertex.label.cex = 0.7, edge.arrow.size = 0.5,
     main = 'Collaboration Network (Authors)', vertex.label = V(g)$name)
#I am not impressed with the plot above. I will try and clean the data and plot again
```

```{r}
# Try installing the stm package if not available
#install.packages('stm', repos='https://cran.rstudio.com/', dependencies=TRUE)

# The code snippet demonstrates the installation of the stm package, processing of text documents, and estimation of topic prevalence over publication years using Structural Topic Modeling.

library(stm)
library(ggplot2)

# Extract summary_findings documents
docs <- df_mergedx$abstract
# Remove empty or NA docs
docs <- docs[!is.na(docs) & docs != ""]

# Create metadata: select publication years corresponding to docs
meta <- data.frame(year = df_mergedx$year[!is.na(df_mergedx$abstract) & df_mergedx$abstract != ""], stringsAsFactors = FALSE)

# Process the documents: tokenization, cleaning, etc.
processed <- textProcessor(documents = docs, metadata = meta)

# Prepare documents for stm
out <- prepDocuments(processed$documents, processed$vocab, processed$meta, lower.thresh = 5)

# Let's assume 5 topics
K <- 5
set.seed(123)
stm_model <- stm(documents = out$documents, vocab = out$vocab, K = K, prevalence = ~year, data = out$meta, max.em.its = 75, init.type = "Spectral")

# Display top words for each topic
labelTopics(stm_model, n = 10)

# Plot topic prevalence
plot(stm_model, type = "summary", main = "Topic Prevalence")

```

```{r}
# Estimate effect of publication year on topics
prep <- estimateEffect(1:K ~ year, stm_model, meta = out$meta, uncertainty = "Global")

# For each topic, plot the effect of year
plot(prep, covariate = "year", topics = 1:5, model = stm_model, method = "continuous", xlab = "Year", main = "Effect of Year on Topics")
```

```{r}
# Create a country collaboration network
library(dplyr)
library(bibliometrix)

# Convert affiliation column to character and split entries
df_processed <- df_mergedx %>%
  mutate(
    affiliation = as.character(affiliation),
    affiliation = gsub("\\s{2,}", " ", affiliation)  # Remove extra spaces
  )

df_processed <- df_processed %>%
  filter(!is.na(affiliation) & affiliation != "")

# If your data is from a bibliographic export (e.g., Web of Science, Scopus)
# df_processed <- convert2df("yourfile.bib", dbsource = "scopus", format = "bibtex")
# # Then run metaTagExtraction
# df_meta <- metaTagExtraction(df_processed, Field = "AU_CO", sep = ";")


```

```{r}
library(stringr)
library(countrycode)
library(purrr)

# Extract country names using regex
df_countries <- df_processed %>%
  mutate(
    countries = str_extract_all(
      affiliation,
      "\\b(?:Germany|United Kingdom|Netherlands|Austria|Norway|Switzerland|United States|Canada|Czech Republic|China|sweden|Spain|Brazil|Hong Kong|Belgium|Chile|Taiwan|Italy|Malawi|Australia|Ecuado|Japan|Ethiopia|Norway|Malayia|Indonesia|South Africa|India|South Korea|France|Morocco|Finlan|Slovenia|Ireland|Kenya|Vietnam|Bangladesh|Nigeria|Demark|Singapore|Uganda|Qatar|Philipines|Pakistan|Burkina Faso|Egypt|Switzerland|Norway|Tanzania)\\b"  # Add all target countries
    )
  ) %>%
  unnest(countries) %>%
  mutate(
    country_code = countrycode(countries, "country.name", "iso3c")  # Standardize names
  )

#####################Creating country pairs for the network analysis################

country_pairs <- df_countries %>%
  group_by(id) %>%  
  summarise(countries = list(unique(country_code))) %>%
  rowwise() %>%  # Process each group individually
  mutate(
    pairs = if (length(countries) > 1) {
      list(combn(countries, 2, simplify = FALSE))  # Generate pairs per group
    } else {
      list(NULL)  # Handle groups with only 1 country
    }
  ) %>%
  unnest(pairs) %>%
  mutate(
    country1 = map_chr(pairs, 1),  # Extract first country in pair
    country2 = map_chr(pairs, 2),   # Extract second country
  ) %>%
  count(country1, country2)  # Count collaborations


# Convert to igraph object
#library(igraph)
g <- graph_from_data_frame(country_pairs, directed = FALSE)
plot(g, layout = layout_with_fr, vertex.label.cex = 0.8)
#add a title: Country Collaboration Network
title("Country Collaboration Network")
# Save the plot
# Save the plot to a file
#ggsave("country_collaboration_network.png", width = 10, height = 8, dpi = 300)


#save plot to directory

```

```{r}


# Degree Centrality
V(g)$degree <- degree(g)

# Betweenness Centrality
V(g)$betweenness <- betweenness(g)

# Closeness Centrality
V(g)$closeness <- closeness(g)

# Community Detection (Louvain)
communities <- cluster_louvain(g)
V(g)$community <- as.factor(membership(communities))

# Edge Density
edge_density(g)

# Diameter
diameter(g)

```

```{r}
library(ggraph)
# Plot degree distribution
ggplot(data.frame(degree = degree(g)), aes(x = degree)) +
  geom_histogram(binwidth = 1, fill = "steelblue") +
  labs(title = "Degree Distribution", x = "Degree", y = "Count")
# Save the plot
#ggsave("degree_distrib.jpg", width = 10, height = 8, dpi = 300)

# Community visualization
ggraph(g, layout = "fr") +
  geom_edge_link(alpha = 0.2) +
  geom_node_point(aes(color = community, size = degree)) +
  theme_void()+
  scale_color_manual(values = rainbow(length(unique(V(g)$community)))) +
  labs(title="Community Visualization", color = "Community", size = "Degree") +
  theme(legend.position = "bottom")

# Save the plot
#ggsave("community_visualization.jpg", width = 10, height = 8, dpi = 300)
```

#Centrality measures for the top five countries

```{r}

# Compute centrality measures
centrality_tbl <- tibble(
  country = V(g)$name,
  degree = degree(g),
  betweenness = betweenness(g),
  closeness = closeness(g)
)

top_countries <- centrality_tbl %>%
  arrange(desc(degree)) %>%
  slice_head(n = 5)

library(knitr)
kable(top_countries, digits = 3, caption = "Top 5 Countries by Degree Centrality")
```

\#*#Next Steps* - Identify the most common visualization types - Identify the most common visualization tools - Identify the key audience or stakeholders for the visualizations - Identify the most common evaluation methods - Prepare a draft report summarizing the findings

```{r}
# Loading the dataset
 data <- read.csv('paper_intext_dta.csv', stringsAsFactors = FALSE)

# For participant types
participant_df <- data %>% 
  filter(!is.na(targets.or.participants)) %>% 
  mutate(participant = strsplit(as.character(targets.or.participants), ",\\s*")) %>% 
  unnest(participant)

participant_counts <- participant_df %>% 
  group_by(participant) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

print(participant_counts)

# Plot with counts on bars 
p6 <- ggplot(participant_counts, aes(x = reorder(participant, count), y = count)) +
  geom_bar(stat = 'identity', fill = 'darkred', width = 0.7) +  # Adjust bar width
  geom_text(aes(label = count), hjust = -0.2, size = 3.5, color = 'black') +  # Add counts
  coord_flip() +  # Horizontal bars
  labs(title = 'Participant Types Counts', x = 'Participant Type', y = 'Count') +
  theme_minimal() +  # Clean background
  theme(
    panel.grid.major.y = element_blank(),  # Remove horizontal grid lines
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(color = "grey90"),  # Light vertical grid
    plot.title = element_text(hjust = 0.5, face = "bold"),  # Centered title
    axis.text = element_text(color = "black"),  # Black axis text
    axis.title = element_text(face = "bold")  # Bold axis titles
  ) +
  expand_limits(y = max(participant_counts$count) * 1.1)  # Add space for labels

print(p6)
```

```{r}
# 1. Types of Data Visualizations
viz_df <- data %>% 
  filter(!is.na(data.visualization.types)) %>% 
  mutate(viz = strsplit(as.character(data.visualization.types), ",\\s*")) %>% 
  unnest(viz)

viz_counts <- viz_df %>% 
  group_by(viz) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

p1 <- ggplot(viz_counts, aes(x = reorder(viz, count), y = count)) +
  geom_bar(stat = 'identity', fill = 'steelblue') +
  coord_flip() +
  labs(title = 'Types of Data Visualizations Used', 
       x = 'Visualization Type', 
       y = 'Count') +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA)
  )

print(p1)

# 2. Tools used for Visualization
tool_df <- data %>% 
  filter(!is.na(data.visualization.tools)) %>% 
  mutate(tool = strsplit(as.character(data.visualization.tools), ",\\s*")) %>% 
  unnest(tool)

tool_counts <- tool_df %>% 
  group_by(tool) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

p2 <- ggplot(tool_counts, aes(x = reorder(tool, count), y = count)) +
  geom_bar(stat = 'identity', fill = 'darkgreen') +
  coord_flip() +
  labs(title = 'Tools Used for Data Visualization', 
       x = 'Tool', 
       y = 'Count') +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA)
  )

print(p2)

# 3. Evaluation Methods
eval_df <- data %>% 
  filter(!is.na(type.of.evaluation)) %>% 
  mutate(eval = strsplit(as.character(type.of.evaluation), ",\\s*")) %>% 
  unnest(eval)

eval_counts <- eval_df %>% 
  group_by(eval) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

p3 <- ggplot(eval_counts, aes(x = reorder(eval, count), y = count)) +
  geom_bar(stat = 'identity', fill = 'purple') +
  coord_flip() +
  labs(title = 'Evaluation Methods Used', 
       x = 'Evaluation Method', 
       y = 'Count') +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA)
  )

print(p3)

# 4. Research Domains
domain_counts <- data %>% 
  filter(!is.na(domain)) %>% 
  group_by(domain) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

p4 <- ggplot(domain_counts, aes(x = reorder(domain, count), y = count)) +
  geom_bar(stat = 'identity', fill = 'orange') +
  coord_flip() +
  labs(title = 'Research Domains', 
       x = 'Domain', 
       y = 'Count') +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA)
  )

print(p4)

# 5. Study Designs
design_df <- data %>% 
  filter(!is.na(study_design)) %>% 
  mutate(design_split = strsplit(as.character(study_design), ",\\s*")) %>% 
  unnest(design_split)

design_counts <- design_df %>% 
  group_by(design_split) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

p5 <- ggplot(design_counts, aes(x = reorder(design_split, count), y = count)) +
  geom_bar(stat = 'identity', fill = 'brown') +
  coord_flip() +
  labs(title = 'Study Designs Used', 
       x = 'Study Design', 
       y = 'Count') +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA)
  )

print(p5)
```

```{r}
# Assuming 'df' is your dataframe with a 'domain' column
# Step 1: Extract and clean domains
domain_counts <- data %>%
  # Drop NA values and convert to character
  drop_na(domain) %>%
  mutate(domain = as.character(domain)) %>%
  # Split comma/separated domains into individual rows
  mutate(domain = strsplit(domain, ",\\s*|;\\s*")) %>%
  unnest(domain) %>%
  # Clean whitespace and convert to lowercase
  mutate(domain = str_trim(tolower(domain))) %>%
  # Filter out empty strings
  filter(domain != "") %>%
  # Count domain frequencies
  count(domain, name = "frequency") %>%
  arrange(desc(frequency))

# Print the most frequent domain
most_freq_domain <- domain_counts$domain[1]
print(paste("Most frequent domain:", most_freq_domain))

# Step 2: Plot (using ggplot2)
ggplot(domain_counts, aes(x = frequency, y = reorder(domain, frequency))) +
  geom_col(fill = "darkmagenta") +  # Bar plot
  geom_text(aes(label = frequency), hjust = -0.2, size = 3.5) +  # Add counts
  labs(
    title = "Frequency of Domains",
    x = "Frequency",
    y = "Domain"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    panel.grid.major.y = element_blank()  # Remove horizontal grid lines
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1)))  # Space for labels
```

```{r}
library(stringr)

# Assume your data frame is called df and the column is study_design
# Remove NA and convert to character
study_series <- data %>%
  filter(!is.na(study_design)) %>%
  mutate(study_design = as.character(study_design))

# Split by comma or semicolon, trim whitespace, and unnest
study_designs_long <- study_series %>%
  mutate(design = str_split(tolower(study_design), "[,;]")) %>%
  unnest(design) %>%
  mutate(design = str_trim(design)) %>%
  filter(design != "")

# Count frequency
design_counts <- study_designs_long %>%
  count(design, sort = TRUE)

# Barplot
ggplot(design_counts, aes(x = n, y = reorder(design, n))) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Frequency of Study Designs", x = "Frequency", y = "Study Design") +
  theme_minimal()

# Most common study design
most_common_design <- design_counts$design[1]
cat("Most common study design:", most_common_design, "\n")

```

```{r}
library(reshape2)

# Get unique designs
unique_designs <- unique(design_counts$design)

# Create a list of designs per study (row)
design_lists <- study_series %>%
  mutate(design = str_split(tolower(study_design), "[,;]")) %>%
  mutate(design = lapply(design, str_trim)) %>%
  mutate(design = lapply(design, function(x) x[x != ""]))

# Initialize co-occurrence matrix
co_mat <- matrix(0, nrow = length(unique_designs), ncol = length(unique_designs),
                 dimnames = list(unique_designs, unique_designs))

# Fill co-occurrence matrix
for (row in design_lists$design) {
  if (length(row) > 1) {
    combos <- combn(row, 2)
    for (i in 1:ncol(combos)) {
      a <- combos[1, i]
      b <- combos[2, i]
      co_mat[a, b] <- co_mat[a, b] + 1
      co_mat[b, a] <- co_mat[b, a] + 1
    }
  }
}

# Convert to data frame for plotting
co_df <- melt(co_mat)
colnames(co_df) <- c("Design1", "Design2", "Count")

# filter out zero-count cells for cleaner labels
co_df_nonzero <- co_df %>% filter(Count > 0)

# Plot heatmap with count labels
ggplot(co_df, aes(x = Design1, y = Design2, fill = Count)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "orangered") +
  geom_text(data = co_df_nonzero, aes(label = Count), color = "black", size = 3) +
  labs(title = "Study Design Co-occurrence Matrix", x = NULL, y = NULL) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

```{r}
cat("Most common study design:", most_common_design, "\n")

```

```{r}
# column is 'type of evaluation'
eval_series <- data %>%
  filter(!is.na(type.of.evaluation)) %>%
  mutate(evaluation_type = as.character(type.of.evaluation))

# Split by comma or semicolon, trim whitespace, and unnest
evals_long <- eval_series %>%
  mutate(eval = str_split(tolower(type.of.evaluation), "[,;]")) %>%
  unnest(eval) %>%
  mutate(eval = str_trim(eval)) %>%
  filter(eval != "")

# Count frequency
eval_counts <- evals_long %>%
  count(eval, sort = TRUE)

# Barplot
ggplot(eval_counts, aes(x = n, y = reorder(eval, n))) +
  geom_bar(stat = "identity", fill = "darkolivegreen3") +
  labs(title = "Frequency of Evaluation Types", x = "Frequency", y = "Evaluation Type") +
  theme_minimal()

```

```{r}
# Get unique evaluation types
unique_evals <- unique(eval_counts$eval)

# Create a list of evals per study (row)
eval_lists <- eval_series %>%
  mutate(eval = str_split(tolower(evaluation_type), "[,;]")) %>%
  mutate(eval = lapply(eval, str_trim)) %>%
  mutate(eval = lapply(eval, function(x) x[x != ""]))

# Initialize co-occurrence matrix
co_mat_eval <- matrix(0, nrow = length(unique_evals), ncol = length(unique_evals),
                      dimnames = list(unique_evals, unique_evals))

# Fill co-occurrence matrix
for (row in eval_lists$eval) {
  if (length(row) > 1) {
    combos <- combn(row, 2)
    for (i in 1:ncol(combos)) {
      a <- combos[1, i]
      b <- combos[2, i]
      co_mat_eval[a, b] <- co_mat_eval[a, b] + 1
      co_mat_eval[b, a] <- co_mat_eval[b, a] + 1
    }
  }
}

# Convert to data frame for plotting
co_df_eval <- melt(co_mat_eval)
colnames(co_df_eval) <- c("Eval1", "Eval2", "Count")
co_df_eval_nonzero <- co_df_eval %>% filter(Count > 0)

# Plot heatmap with count labels
ggplot(co_df_eval, aes(x = Eval1, y = Eval2, fill = Count)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "darkolivegreen") +
  geom_text(data = co_df_eval_nonzero, aes(label = Count), color = "black", size = 3) +
  labs(title = "Evaluation Type Co-occurrence Matrix", x = NULL, y = NULL) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
